Finding a closest point in a lattice of Voronoi's first kind

### 2014-07-12 reviews/response.tex

%The authors feel the paper has improved as a result.

\section*{Reviewer 1}\label{sec:reviewer-1}

\begin{itemize}

\item\textbf{Comment:} 
From the perspective of quadratic forms, the quadratic forms induced by obtuse superbases are in one to one correspondence with graph Laplacians (this is not
explicitly stated in the paper for some reason). Indeed, the quadratic form $\|B\xbf\|^2$ is exactly $\xbf^T L_G \xbf$, where $L_G$ is the Laplacian of the weighted graph $G$ on vertices $\{ 1, . . . , n + 1 \}$ , with edges weights $w_{ij} = -\bbf_i \cdot \bbf_j \geq 0, i = j$ (which they call the Selling parameters). More precisely:
\[
\|B\xbf\|_2^2 = \xbf^T L_G \xbf = \sum_{ \{i,j\} \in E[G] } w_{ij} (\xbf_{i} - \xbf_{j} )^2.
\]
\\ \textbf{Response:} We thank the reviewer for pointing out the connection with the Laplacian matrix of a graph.  The following text has been included in Section~\zref{sec:comp-clos-relev} to highlight this link.

@@

\end{itemize}

We now respond to the reviewers numbered comments.

\begin{enumerate}

\item\textbf{Comment:} 
I believe Proposition 2.1 and 2.2 appear essentially verbatim in [8]. Same for Propositions
3.1 (which is essentially by definition) and 3.2, they appear in [9]. I don’t think it adds any
readability/anything of value to reprove them here.
\\\textbf{Response:}

@@

We agree that these propositions are well known/straightforward, but would prefer to keep them so that the paper is self contained. We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).  The proof of Proposition 2.1 has been shortened in response to comment~\ref{com:rev2halfopen} of reviewer 2.  A result like Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.  As far as we are aware, the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume).  Instead, the procedure operates recursively on scaled versions of the lattice.  The scaling is chosen to ensure that the target point always lies inside twice the Voronoi cell.

Motivating Propositions 3.1 and 3.2 is that the stopping criterion in our procedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.  The stopping criteria in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013} essentially involve a direct check that a given point, say $x$, is contained inside the Voronoi cell.  That is, one directly asserts that $2 x \cdot v \leq \|v\|^2$ for all relevant vectors $v$.  We cannot use this approach since it would amount to checking this inequality over as many as $2^{n+1}-2$ relevant vectors.  Instead we simply check that $\|y - x_K\| \leq \|y - x_{K+1}\|$.  That is, we check that the procedure converges in the sense that 
\[
\|y - x_K\| = \|y - x_{K+1}\| = \|y - x_{K+2}\| = \dots.
\]
This is a straightforward, but important, difference and it motivates Propositions 3.1 and 3.2.

@@

\newcommand{\rng}{\operatorname{rng}}
\newcommand{\subrng}{\operatorname{subr}}
\newcommand{\decrng}{\operatorname{decrng}}

\item\textbf{Comment:} \label{com:uselaplaicesubr1}
Almost all the inequalities in the paper can be derived from the following simple observation.  For an $n$ vertex weighted graph $G$ with non-negative edge weights, and vectors, $\xbf,\ybf \in \reals^n$, if there exists a permutation $\pi$ of $[n]$ such that $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \leq \dots \leq \ybf_{\pi[n]}$ then
\[
\xbf^T L_G \ybf = \sum_{ \{i,j\} \in E[G] } \wbf_{ij} (\xbf_{i} - \xbf_{j}) (\ybf_{i} - \ybf_{j}) \geq 0.
\]
Furthermore, if $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \geq \dots \geq \ybf_{\pi[n]}$ then the above inequality if reversed, i.e $\xbf^T L_G \ybf \leq 0$.

I believe using the above will make many of your proofs more transparent.  For example, if you redefine $\subrng(\xbf)$ to return the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq 1$, then notice that if $S \neq \emptyset$ then there is a permutation that puts both $\xbf - 1_S$ and $1_S$ in non-decreasing order. In particular this implies that
\[
\xbf L_G \xbf = (\xbf - 1_S)^T L_G (\xbf - 1_S) + 2(\xbf - 1_S)^T L_G 1_S + 1_S^T L_G 1_S \geq (\xbf - 1_S)^T L_G (\xbf - 1_S),
\]
i.e. $\xbf$ can be made ``shorter'' by squeezing its components closer together.
\\\textbf{Response:}

The property described by the reviewer is indeed needed to give a proof of Theorem 4.1 (see our response to the next comment number~\ref{com:provethm4.1}).  However, this is not the property that is needed to prove that our algorithm converges in at most $n$ steps.  To see this, let $\subrng_k(p)$ denote the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq k$. Thus $\subrng_1$ is that suggested by the reviewer and $\subrng_2$ is that which is used in the paper.  Our proof works by considering a $v \in \ints^n$ that can be added to $\floor{z}$ so that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$.  Our Lemma~\zref{lem:roundzclose} shows that there exists such a $v$ with $\subrng_2(v) = \emptyset$ and so $\rng(v) \leq n$.  It is not true that there always exists such a $v$ with  $\subrng_1(v) = \emptyset$.

To consider this in more detail it is worth looking at the proof of Lemma~\zref{lem:decrngpreservesclosestpoints}.  The lemma shows that $B(\floor{z} + v)$ is a closest point to $y = Bz$ only if $B\big(\floor{z} + \decrng(v)\big)$ is.  A key part of the proof is the line
\[
1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0 \qquad \text{for $i \in S$ and $j \notin S$}
\]
where $\zeta_i = z_i - \floor{z_i} \in [0,1)$.  Since $\zeta_i-\zeta_j \in (-1,1)$ we require that the gap between $v_i - v_j$ to be greater than $\tfrac{3}{2}$ for this inequality to hold.  Because $v \in \ints^{n+1}$ this means $v_i - v_j \geq 2$.  We cannot replace the current $\subrng_2$ with the suggested $\subrng_1$ and have this proof hold because then we would only have $v_i - v_j \geq 1$.

@@

% Very early on (more than 2 years ago now!) when first looking at lattices of Voronoi's first kind, the first author believed that a single iteration of the proceedure would surfice.  That is, compute
% \[
% v = \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - \floor{z} - t) \|^2
% \]
% and then $B(\floor{z} + v)$ would be a closest lattice point.  Numerical counterexamples dispelled this possibility.

\item\textbf{Comment:} \label{com:provethm4.1}
Furthermore, one can use this to give a two line proof of Theorem 4.1.  This theorem I actually do think make sense to reprove here, because modifications of this proof are what drives the convergence bound. 
\\\textbf{Response:}

The property described by the reviewer is indeed needed to give a proof of Theorem 4.1.  To our knowledge, one also needs the property that the relevant vectors of a lattice are given by the shortest coset representatives of $\Lambda / 2\Lambda$.  This result is attributed to Voronoi (see for example \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} and \cite[Theorem~3.2]{MicciancioVoulgaris_deterministic_jv_2013}).  This is used by Conway and Sloane in~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.

We believe that the proof, whilst not difficult, is ultimately a distraction from our main results.  This would be different if properties needed for the proof, such as the function $\subrng_1$ and Selling's formula were also needed for proving the convergence of our algorithm.  However, as we have stated in comment~\ref{com:uselaplaicesubr1} above, Selling's formula is not directly needed, and $\subrng_2$, not $\subrng_1$ is the function required.

The order of some paragraphs and lemmas has been changed in Section~\zref{sec:seri-relev-vect} with the intent of improving the readability of the proofs and presenting the intuition behind the proofs before formalities begin.

@@

\item\textbf{Comment:} 
Lemma 5.7: $h$ is ill-defined if $\rng(\vbf) = 0$.  You should explicitly state that $\rng(\vbf) \geq 1$.
\\\textbf{Response:}
Fixed.

\end{enumerate}

\section*{Reviewer 2}\label{sec:reviewer-2}

\begin{enumerate}
\item\textbf{Comment:}  \label{com:cvvp}
page 1, last line ``The closest lattice point problem is known to be NP-hard under certain condition ...'': I found this paragraph very confusing, and somehow inaccurate. What ``certain conditions''? I think a better way to describe the situation regarding the NP-hardness of CVP in relation to this paper is the following. CVP is known to be NP-hard even to approximate within almost polynomial factor (see 35) when the lattice is arbitrary and not known in advance. Interestingly (see 35 and 17) there are families of lattices when CVP remain NP-hard even when given the best possible basis (or any hard-to-compute side information about the lattice). This version of the problem is called CVP with preprocessing (CVPP), and it is particularly relevant to this work, where it is shown that CVPP can be solved exactly and in polynomial time for the class of lattices of Voronoi first kind, when given as auxiliary information a special basis (namely, an obtuse superbasis.) 
\\\textbf{Response:}

@@

\item\textbf{Comment:}  
page 2, second paragraph: Aren't the results of 3,29 and 30 subsumed by 34? (The authors sure know better than me. A a short clarification may direct the interested reader to the best paper to read.) Also, CVP is always solvable in polynomial time when the dimension is fixed. So, it is unclear the significance of CVP algorithms for Leech lattice (in fixed dimension 24) in the context of this discussion. 
\\\textbf{Response:}
This sentence has been reworded and the reference list has been reduced.  It now reads:

``Although the problem is NP-hard in general, fast algorithms are known for specific highly regular lattices, such the integer lattice $\ints^{n}$, the root lattices $A_n$ and $D_n$, their dual lattices $A_n^*$ and $D_n^*$, and the related Coxeter lattices~\cite[Chap.~4]{SPLAG}\cite{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices}.''

The interested reader could read these references in the order given. %, i.e., first Chapter 4 of~\cite{SPLAG} and~\cite{Conway1982FastQuantDec} for $D_n$ and $D_n^*$ and older algorithms for $A_n$ and $A_n^*$.  Then~\cite{McKilliam2008} for a fast algorithm for $A_n^*$, then~\cite{McKilliam2009CoxeterLattices} for fast algorithms for all of the Coxeter lattices that includes $A_n$ and $A_n^*$, $E_8$, etc.  Technically, the algorithm from~\cite{McKilliam2009CoxeterLattices} does include that from~\cite{McKilliam2008}.  However~\cite{McKilliam2008} is simpler and shorter to read first.

\item\textbf{Comment:}  
When saying ``if the obtuse superbasis is known'', you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 
\\\textbf{Response:}
Please see our response to comment~\ref{com:cvvp}.

\item\textbf{Comment:}  
page 2,bottom, displayed equation ``$v \cdot x \leq x \cdot x$''. Maybe you mean strict inequality.
\\\textbf{Response:}
%In our original definition we intended to include what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  This we not necessary
Yes, this has been fixed.  The text now reads reads:

@@

% Our definition includes what Conway and Sloane call the `lax' relevant vectors.  To make this clear the following sentences have been added below the definition of the relevant vectors in Section~\zref{sec:voron-cells-relev}:

% ``This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.'' %It will be of benefit to include all of the strict and lax relevant vectors here.''

\item\textbf{Comment:}  
page 3, and maybe elsewhere ``short vector in a lattice'' should probably be ``shortest nonzero vector in a lattice''. 
\\\textbf{Response:}
We have instead introduced the term `short vector' in the introduction.  The following sentence is included in the first paragraph of the introduction: ``This is called the \emph{shortest vector problem} and a solution is called a~\emph{short vector}.''

\item\textbf{Comment:}  
page 3, definition of ceiling(r) as smallest integer strictly larger than 3 seems unconventional. The standard definition is not strict. Write floor(r)+1 instead, which is consistent with the definition of floor(r) given later in the paper. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  \label{com:rev2halfopen}
page 3, proof of Prop. 2.1. Your definition of ``half-open Voronoi cell'' is not a proper definition, and also seems incorrect. E.g., take the integer lattice $\ints^n$, then only one of the corners (rather than half) of the corners of the Voronoi cell should be included if you want a tiling. 
\\\textbf{Response:}
The reviewer is correct.  We only require that some modification of the boundary of the closed Voronoi cell results in a tiling.  The proof of Proposition~\zref{eq:latticepointsinvorcvell} has been modified accordingly.

@@

\item\textbf{Comment:}  
page 3, prop 2.2, again definition of ceiling(r) is non-standard. Use floor(r)+1 instead. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  
page 3, section 3. 
The process described in (3.1) is almost identical to those in [45,37], so those papers should probably be referenced at the very beginning, and a comparison provided. As far as I remember, the difference is that 37 picks the v that minimizes the Voronoi norm (i.e., the norm defined by the Voronoi cell of the origin) instead of the Euclidean norm. On the other hand, 45 uses the Euclidean norm, but selects an arbitrary relevant vector that reduces the Euclidean norm, rather than the one that reduces it the most. In this respect, your process is a special case of the one described in 45, and termination (your propositions 3.1 and 3.2) is already proved in 45. 
\\\textbf{Response:}
We have included the following text in the first paragraph of Section~\zref{sec:iterative-slicer} when introducing the iterative algorithm.

@@

``Motivated by Shalvi~et.~al.~\cite{Shalvi_iterativeslicer_2009} and  Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013}, we consider the following iteration,''

Among the iterative procedures in our paper, in~\cite{Shalvi_iterativeslicer_2009}, and in~\cite{MicciancioVoulgaris_deterministic_jv_2013}, ours is the simplest.  The choice of our iteration is not motivated by what might be fastest in general lattices.  It is motivated by what we can prove to be polynomial time for those lattices of Voronoi's first kind.

@@

The~\emph{iterative slicer}~\cite{Shalvi_iterativeslicer_2009} is similar but uses the (presumably accelerated?) iteration
\begin{align*}
x_{k+1} &= x_k + \alpha_k v_k  \\
(v_k, \alpha_k) &= \arg\min_{ (v, \alpha) \in ( \relevant(\Lambda) \cup \{\zerobf\}, \ints) } \|y - x_k - \alpha v \|.
\end{align*} 
So it allows adding integer multiples of a relevant vector.  We do not believe using this will improve the complexity of our algorithm beyond $O(n^4)$.  It would greatly complicate our proofs.

The iterative algorithm of Micciancio and Voulgaris~\cite{MicciancioVoulgaris_deterministic_jv_2013} is different again.  To our (admittedly rough) knowledge, the procedure computes an $\alpha_k$ (different, but related to that above) and then proceeds to compute a closest point in the scaled lattice $\alpha_k \Lambda$.  The scaling $\alpha_k$ is chosen so that the target point lies within $2\alpha_k \vor(\Lambda)$.  Once a closest lattice point in $\alpha_k\Lambda$ is found, the next $\alpha_{k+1}$ is computed and the procedure repeats.

We believe that detailing the differences between these iterative procedures would only detract from the core purpose of our paper to describe an $O(n^4)$ time algorithm to compute a closest point in lattice of Voronoi's first kind.

@@

\item\textbf{Comment:}  
page 5, as before, you should probably mention also Voulgaris when citing 37. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  
page 5, last line, ``dimension less than 4'', does it include 4 or not? 
\\\textbf{Response:}
It does not include 4.  We have modified the text here to read ``less than or equal to $3$''. We have also modified similar text in the last paragraph of Section~\zref{sec:discussion} to read ``It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.  Lattices of dimension larger than 3 need not be of Voronoi's first kind.''

@@

\item\textbf{Comment:}  
page 5, as before, you should probably mention also Voulgaris when citing 37. 
\\\textbf{Response:}
Done.

\item\textbf{Comment:}  
page 6, Theorem 4.1 (this follows from [ 9, Theorem 2+3], not just theorem 3. 
Also, make the theorem statement self contained by adding ``Let $L$ be a lattice of Voronoi first kind with obtuse superbasis $b_1,\dots,b_{n+1}$''. 
\\\textbf{Response:}
We have included the text ``Let $\Lambda$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.'' into the statement of Theorem 4.1.

### 2014-07-12 paper.tex
added changes

%We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.  
Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.  We want to find $n$ integers $w_1,\dots,w_n$ that minimise
\[
\| y - \sum_{i=1}^n b_i w_i \|^2.
\]
We can equivalently find $n+1$ integers $w_1,\dots,w_{n+1}$ that minimise
\[
\| y - \sum_{i=1}^{n+1} b_i w_i \|^2.
\]
The iterative procedure described in~\eqref{eq:relvectsminimslicer} will be used to do this.  In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.  This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.  Let
\[
B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
\]
be the $m$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.  We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
\begin{equation}\label{eq:tominimise}
\| B(z  -  w) \|^2
\end{equation}
is minimised.  Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ operates on vectors elementwise. In view of Theorem~\ref{thm:revvecssuperbase} the iterative procedure~\eqref{eq:relvectsminimslicer} to compute a closest lattice point can be written in the form
\begin{align}
x_{k+1} &= B u_{k+1} \label{eq:xseqfirsttype}  \\
u_{k+1} &= u_k + t_k \nonumber \\
t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2, \label{eq:pvecmin}
\end{align}
where $\{0,1\}^{n+1}$ denotes the set of column vectors of length $n+1$ with elements equal to zero or one.  The procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.  This choice of initial lattice point is important.  In Section~\ref{sec:comp-clos-relev} we show how minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.  The minimiser may not be unique corresponding with the existence of multiple minimum cuts.  In this case any one of the minimisers may be chosen.  Our results do not depend on this choice.   In the remainder of this section we prove that this iterative procedure results in a closest lattice point after at most $n$ iterations.  That is, we show that there exists a positive integer $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.

@@

We now proceed with a formal proof culminating in Theorem~\ref{thm:mainitrconvthm} stated at the end of this section.  It is necessary to first introduce some notation.  For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and zero otherwise. %Thus, for example, if $p \in \reals^{n+1}$, then $p+\onebf_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.  
For $S \subseteq \{1,\dots,n+1\}$ and $p \in \reals^{n+1}$ define the function
\[
\Phi(S, p) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)
\]
where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters from~\eqref{eq:obtusecond}.  Denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i \in \{1,\dots,n+1\} \mid i \notin S\}$.

@@

\begin{theorem}\label{thm:mainitrconvthm}
Let $\Lambda \subset \reals^m$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.  Let $y \in \reals^m$ and let $x_0,x_1,\dots$ be a sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.  There exists $K \leq n$ such that $x_K$ is a closest lattice point to $y$.
\end{theorem}
\begin{proof}
Recall that we, without loss of generality, suppose that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$ and so $y = Bz$ where $z \in \ints^{n+1}$.  Let $x_k = B u_k$ be the lattice point obtained on the $k$th iteration of the procedure.  Suppose that $x_k$ is $\ell$-close to $y$ with $\ell > 0$.  The procedure computes $t_{k} \in \{0,1\}^{n+1}$ satisfying
\[
\|B(z - u_k - t_{k})\|^2 = \min_{t \in \{0,1\}^{n+1}}\|B(z - u_k - t)\|^2
\]
and puts $x_{k+1} = B(u_k + t_k)$.  It follows from Lemma~\ref{lem:rngdecreases} that $x_{k+1}$ is $(\ell-1)$-close to $y$.  By Lemma~\ref{lem:roundzclose} the lattice point that initialises the procedure $x_0 = B \floor{z}$ is $K$-close to $y$ where $K \leq n$.  Thus, $x_1$ is $(K-1)$-close, $x_2$ is $(K-2)$-close, and so on until $x_K$ is $0$-close.  That $x_K$ is a closest lattice point to $y$ follows from Lemma~\ref{lem:rngzeroclosestpoint}.
\end{proof}

@@

Let $G$ be an undirected graph with $n+3$ vertices $v_0, \dots, v_{n+2}$ contained in the set $V$ and edges $e_{ij}$ connecting $v_i$ to $v_j$.  To each edge we assign a \emph{weight} $w_{ij} \in \reals$.  The graph is undirected so the weights are symmetric, that is, $w_{ij} = w_{ji}$.  By calling the vertex $v_0$ the \emph{source} and the vertex $v_{n+2}$ the \emph{sink} the graph $G$ is what is called a \emph{flow network}.  The flow network is \emph{undirected} because the weights assigned to each edge are undirected.  A \emph{cut} in the flow network $G$ is a subset $C \subset V$ of vertices with its complement $\bar{C} \subset V$ such that the source vertex $v_0 \in C$ and the sink vertex $v_{n+2} \in \bar{C}$.  %The vertex $v_0$ is typically called the \emph{source} and the vertex $v_{n+1}$ is typically called the \emph{sink}.

The weight of a cut is
\[
W(C,\bar{C}) = \sum_{i \in I} \sum_{j \in J} w_{ij}, 
\]
where $I = \{ i \mid v_i \in C\}$ and $J = \{j \mid v_j \in \bar{C}\}$.  That is, $W(C,\bar{C})$ is the sum of the weights on the edges crossing from the vertices in $C$ to the vertices in $\bar{C}$.  In what follows we will often drop the argument and write $W$ rather than $W(C,\bar{C})$.  A \emph{minimum cut} is a $C$ and $\bar{C}$ that minimise the weight $W$.  If all of the edge weights $w_{ij}$ for $i \neq j$ are nonnegative, a minimum cut can be computed in 
order $O(n^3)$ arithmetic operations~\cite{Cormen2001,Even_graph_algorithms_1979}.

We require some properties of the weights $w_{ij}$ in relation to $W$.  If the graph is allowed to contain loops, that is, edges from a vertex to itself, then the weight of these edges $w_{ii}$ has no effect on the weight of any cut.  We may choose any values for the $w_{ii}$ without affecting $W$.  We will find it convenient to set $w_{0,0} = w_{n+2,n+2} = 0$.  The remaining $w_{ii}$ we shall specify shortly.  The edge $e_{0,n+2}$ is in every cut.  If a constant is added to the weight of this edge, that is, $w_{0,n+2}$ is replaced by $w_{0,n+2} + c$ then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.  In particular, the subsets $C$ and $\bar{C}$ corresponding to a minimum cut are not changed.  We will find it convenient to choose $w_{0,n+2} = w_{n+2,0} = 0$.

>>>>

%We are interested in solving the closest lattice point problem for lattices of Voronoi's first kind.  
Let $\Lambda \subset \reals^m$ be an $n$ dimensional lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$ and let $y \in \reals^m$.  We want to find $n$ integers $w_1,\dots,w_n$ that minimise
\[
\| y - \sum_{i=1}^n b_i w_i \|^2.
\]
We can equivalently find $n+1$ integers $w_1,\dots,w_{n+1}$ that minimise
\[
\| y - \sum_{i=1}^{n+1} b_i w_i \|^2.
\]
The iterative procedure described in~\eqref{eq:relvectsminimslicer} will be used to do this.  In what follows it is assumed that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$.  This assumption is without loss of generality because $x$ is a closest lattice point to $y$ if and only if $x$ is a closest lattice point to the orthogonal projection of $y$ into the space spanned by $b_1,\dots,b_{n}$.  Let
\[
B = (b_1\,\,b_2\,\,\dots\,\,b_{n+1})
\]
be the $m$ by $n+1$ matrix with columns given by $b_1,\dots,b_{n+1}$ and let $z \in \reals^{n+1}$ be a column vector such that $y = Bz$.  We now want to find a column vector $w = (w_1,\dots,w_{n+1})^\prime$ of integers such that
\begin{equation}\label{eq:tominimise}
\| B(z  -  w) \|^2
\end{equation}
is minimised.  Define the column vector $u_0 = \floor{z}$ where $\floor{\cdot}$ operates on vectors elementwise. In view of Theorem~\ref{thm:revvecssuperbase} the iterative procedure~\eqref{eq:relvectsminimslicer} to compute a closest lattice point can be written in the form
\begin{align}
x_{k+1} &= B u_{k+1} \label{eq:xseqfirsttype}  \\
u_{k+1} &= u_k + t_k \nonumber \\
t_k &= \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - u_k - t) \|^2 \label{eq:pvecmin}
\end{align}
where $\{0,1\}^{n+1}$ denotes the set of column vectors of length $n+1$ with elements equal to zero or one.  The procedure is initialised at the lattice point $x_0 = Bu_0 = B\floor{z}$.  This choice of initial lattice point is important.  In Section~\ref{sec:comp-clos-relev} we show how minimisation over $\{0,1\}^{n+1}$ in~\eqref{eq:pvecmin} can be computed efficiently in $O(n^3)$ operations by computing a minimum cut in an undirected flow network.  The minimiser may not be unique corresponding with the existence of multiple minimum cuts.  In this case any one of the minimisers may be chosen.  Our results do not depend on this choice.   In the remainder of this section we prove that this iterative procedure results in a closest lattice point after at most $n$ iterations.  That is, we show that there exists a positive integer $K \leq n$ such that $x_K$ is a closest lattice point to $y = Bz$.

@@

We now proceed with a formal proof culminating in Theorem~\ref{thm:mainitrconvthm} stated at the end of this section.  We first introduce some notation.  For $S$ a subset of indices $\{1,\dots,n+1\}$ let $\onebf_S$ denote the column vector of length $n+1$ with $i$th element equal to one if $i \in S$ and zero otherwise. %Thus, for example, if $p \in \reals^{n+1}$, then $p+\onebf_S$ denotes the vector $p$ with those elements having index from $S$ incremented by one.  
For $S \subseteq \{1,\dots,n+1\}$ and $p \in \reals^{n+1}$ define the function
\[
\Phi(S, p) = \sum_{i \in S}\sum_{j \notin S}q_{ij}(1 + 2p_i - 2p_j)
\]
where $q_{ij} = \dotprod{b_i}{b_j}$ are the Selling parameters from~\eqref{eq:obtusecond}.  Denote by $\bar{S}$ the complement of the set of indices $S$, that is $\bar{S} = \{ i \in \{1,\dots,n+1\} \mid i \notin S\}$.

@@

\begin{theorem}\label{thm:mainitrconvthm}
Let $\Lambda \subset \reals^m$ be a lattice of Voronoi's first kind with obtuse superbasis $b_1,\dots,b_{n+1}$.  Let $z \in \reals^{n+1}$, let $y = Bz$, and let $x_0,x_1,\dots$ be a sequence of lattice points given by the iterative procedure~\eqref{eq:xseqfirsttype}.  There exists $K \leq n$ such that $x_K$ is a closest lattice point to $y$.
\end{theorem}
\begin{proof}
%Recall that we suppose, without loss of generality, that $y$ lies in the space spanned by the basis vectors $b_1,\dots,b_{n}$ and so $y = Bz$ where $z \in \ints^{n+1}$.  
Let $x_k = B u_k$ be the lattice point obtained on the $k$th iteration of the procedure.  Suppose that $x_k$ is $\ell$-close to $y$ with $\ell > 0$.  The procedure computes $t_{k} \in \{0,1\}^{n+1}$ satisfying
\[
\|B(z - u_k - t_{k})\|^2 = \min_{t \in \{0,1\}^{n+1}}\|B(z - u_k - t)\|^2
\]
and puts $x_{k+1} = B(u_k + t_k)$.  It follows from Lemma~\ref{lem:rngdecreases} that $x_{k+1}$ is $(\ell-1)$-close to $y$.  By Lemma~\ref{lem:roundzclose} the lattice point that initialises the procedure $x_0 = B \floor{z}$ is $K$-close to $y$ where $K \leq n$.  Thus, $x_1$ is $(K-1)$-close, $x_2$ is $(K-2)$-close, and so on until $x_K$ is $0$-close.  That $x_K$ is a closest lattice point to $y$ follows from Lemma~\ref{lem:rngzeroclosestpoint}.
\end{proof}

@@

Let $G$ be an undirected graph with $n+3$ vertices $v_0, \dots, v_{n+2}$ contained in the set $V$ and edges $e_{ij}$ connecting $v_i$ to $v_j$.  To each edge we assign a \emph{weight} $w_{ij} \in \reals$.  The graph is undirected so the weights are symmetric, that is, $w_{ij} = w_{ji}$.  By calling the vertex $v_0$ the \emph{source} and the vertex $v_{n+2}$ the \emph{sink} the graph $G$ is what is called a \emph{flow network}.  The flow network is undirected because the weights assigned to each edge are undirected.  A \emph{cut} in the flow network $G$ is a subset $C \subset V$ of vertices with its complement $\bar{C} \subset V$ such that the source vertex $v_0 \in C$ and the sink vertex $v_{n+2} \in \bar{C}$.  %The vertex $v_0$ is typically called the \emph{source} and the vertex $v_{n+1}$ is typically called the \emph{sink}.

The weight of a cut is
\[
W(C,\bar{C}) = \sum_{i \in I} \sum_{j \in J} w_{ij}, 
\]
where $I = \{ i \mid v_i \in C\}$ and $J = \{j \mid v_j \in \bar{C}\}$.  That is, $W(C,\bar{C})$ is the sum of the weights on the edges crossing from the vertices in $C$ to the vertices in $\bar{C}$.  In what follows we often drop the argument and write $W$ rather than $W(C,\bar{C})$.  A \emph{minimum cut} is a $C$ and $\bar{C}$ that minimise the weight $W$.  If all of the edge weights $w_{ij}$ for $i \neq j$ are nonnegative, a minimum cut can be computed in 
order $O(n^3)$ arithmetic operations~\cite{Cormen2001,Even_graph_algorithms_1979}.

We require some properties of the weights $w_{ij}$ in relation to $W$.  If the graph is allowed to contain loops, that is, edges from a vertex to itself, then the weight of these edges $w_{ii}$ has no effect on the weight of any cut.  We may choose any values for the $w_{ii}$ without affecting $W$.  It will be convenient to set $w_{0,0} = w_{n+2,n+2} = 0$.  The remaining $w_{ii}$ shall be specified shortly.  The edge $e_{0,n+2}$ is in every cut.  If a constant is added to the weight of this edge, that is, $w_{0,n+2}$ is replaced by $w_{0,n+2} + c$ then $W$ is replaced by $W + c$ for every $C$ and $\bar{C}$.  In particular, the subsets $C$ and $\bar{C}$ corresponding to a minimum cut are not changed.  It will be convenient to choose $w_{0,n+2} = w_{n+2,0} = 0$.
